---
title: "EDA Project Part 3"
author: "Data 350"
date: "Due date: May 4 2026 at midnight (EST)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(shiny)
library(dplyr)
library(purrr)
library(ggplot2)
library(sf)
library(spdep)
library(tidycensus)
library(scales)
library(tibble)
library(tidyr)
library(leaflet)

options(tigris_use_cache = TRUE, sf_max.plot.default = 16^6)
```

```{r}
lisa_quadrants <- function(x, w, alpha = 0.05) {
  n <- length(x)
  out <- rep("No data", n)
  ok <- !is.na(x)
  if (!any(ok) || sum(ok) < 2) {
    return(out)
  }
  z <- rep(NA_real_, n)
  z[ok] <- as.numeric(scale(x[ok]))
  lag_z <- spdep::lag.listw(w, z, zero.policy = TRUE)
  lm <- spdep::localmoran(x, w, zero.policy = TRUE)
  p <- lm[, "Pr(z != E(Ii))"]
  dplyr::case_when(
    !ok ~ "No data",
    is.na(p) | is.na(z) | is.na(lag_z) ~ "No data",
    p > alpha ~ "Not significant",
    z > 0 & lag_z > 0 ~ "High–High (high inequality cluster)",
    z < 0 & lag_z < 0 ~ "Low–Low",
    z > 0 & lag_z < 0 ~ "High–Low (outlier)",
    z < 0 & lag_z > 0 ~ "Low–High (outlier)",
    TRUE ~ "Other"
  )
}

build_county_weights <- function(geom_sf) {
  geom <- sf::st_geometry(geom_sf)
  nb <- spdep::poly2nb(geom, queen = TRUE)
  spdep::nb2listw(nb, style = "W", zero.policy = TRUE)
}

fetch_gini_year <- function(state, year, census_key) {
  tidycensus::get_acs(
    geography = "county",
    variables = "B19083_001",
    state = state,
    year = year,
    survey = "acs5",
    geometry = TRUE,
    key = census_key
  ) |>
    dplyr::rename(gini = estimate, gini_moe = moe) |>
    dplyr::mutate(
      year = year,
      county_name = NAME
    ) |>
    dplyr::filter(!is.na(gini))
}

ui <- fluidPage(
  tags$head(tags$style(HTML("
    body { font-family: 'Helvetica Neue', Helvetica, Arial, sans-serif; background: #f8f7f4; color: #2c2c2a; }
    .well { background: #fff; border: 1px solid #e0ddd4; border-radius: 10px; box-shadow: none; }
    .nav-tabs { border-bottom: 2px solid #e0ddd4; }
    .nav-tabs > li.active > a { font-weight: 600; color: #2c2c2a; }
    .section-label { font-size: 11px; color: #888780; text-transform: uppercase; letter-spacing: .06em; margin: 14px 0 4px; }
    .guiding-q { background: #1a2e4a; color: #fff; padding: 18px 28px; border-radius: 10px; margin-bottom: 20px; }
    .guiding-q h3 { margin: 0; font-size: 20px; font-weight: 500; }
    .guiding-q p { margin: 6px 0 0; font-size: 13px; color: #a0b0c8; }
  "))),

  div(class = "guiding-q",
    h3("How does socioeconomic inequality vary across geographic areas, and what factors are most strongly associated with it?")
  ),

  tabsetPanel(
    id = "main_tabs",

    # ── Tab 1: Inequality Clusters ──────────────────────────────────────────
    tabPanel(
      "Inequality Clusters",
      br(),
      sidebarLayout(
        sidebarPanel(
          width = 3,
          div(class = "section-label", "Question"),
          p("Are there clusters of high inequality, and how have they changed over time?",
            style = "font-size:12px; color:#6b7a90;"),
          hr(),
          div(class = "section-label", "Map Settings"),
          sliderInput("year_map", "Map Year",
            min = 2010L, max = 2024L, value = 2019L, step = 1L, sep = ""),
          radioButtons("map_fill", "Fill counties using",
            choices = c("Gini index" = "gini", "LISA cluster type" = "lisa"),
            selected = "gini", inline = FALSE),
          hr(),
          actionButton("fetch", "Load Data", class = "btn-primary", style = "width:100%;")
        ),
        mainPanel(
          width = 9,
          uiOutput("map_or_placeholder"),
          br(),
          div(
            style = "background:#fff; border:1px solid #e0ddd4; border-radius:10px; padding:22px 28px; margin-top:10px;",
            h4("Analysis", style = "font-size:15px; font-weight:600; margin-top:0; color:#2c2c2a;"),
            p("This is map displaying levels of income inequality in Maryland. To measure income inequality, we use two metrics: Gini index and LISA cluster. Gini index is a measure of income inequality ranging from 0 (perfect equality) to 1 (maximum inequality). LISA (Local Indicators of Spacial Association) clusters are meant to detect \"hot spots\" and \"cold spots\", which are meant to identify groupings of high or low income inequality."),
            p("When Gini index is selected, counties are colored based on their Gini index, with darker/warmer tones indicating higher inequality. However, there is also a margin of error associated with each county, which is reflected in their opacity - a lower transparency indicates a higher margin of error, and as such, should be treated with more caution when interpreting this visualization. Finally, the borders of each county, while somewhat difficult to see, indicate LISA cluster types."),
            p("When LISA cluster is selected, counties are instead colored based on their LISA (Local Indicators of Spatial Association) cluster, which was computerd using Anselin's Local Moran's I statistic (with alpha = 0.05). Red (High-High) counties have high inequality and are surrounded by high-inequality neighbors, blue (Low-Low) counties have low inequality and are surrounded by low-inequality neighbors. Additionally, orange and purple counties are outliers (where their inequality level diverges sharply from their neighbors), and gray counties are not statistically significant."),
            p("Comparing Gini index throughout time, earlier years appear to have more uncertainty, as the coloring is more transparent. This is likely a result of the data - these earlier years may not have had as strict or careful data collection, resulting in higher margins of error. Moving through time, there are a few years where Gini index changes - between 2013 and 2014, and in the past ~5 years. However, there does not appear to be a strong trend in which direction Gini index changes (towards more or less inequality), or which counties are changing the most. Instead, we can see that more urban counties between Washington DC and Baltimore have lower levels of income inequality, while more rural counties and Baltimore City have higher levels of income inequality."),
            p("Comparing LISA clusters throughout time, between 2010-2017, the only counties with statistically significant clusters are in Southern Maryland (Charles, Calvert, and St. Mary's Counties). However, between 2018-2019, Queen Anne's County was also labelled as low-low inequality, and in 2022, Prince George's County was also considered to have low-low inequality. This lack of statistical significance or change in which counties are labelled indicates that there are no strong clusters of inequality (geographic areas of high-high or low-low inequality."),
            p("In conclusion, this map helps us visualize income inequality geographically in Maryland. Namely, there are lower levels of income inequality in more urban counties between Washington DC and Baltimore, while there are higher levels of income inequality in more rural counties and Baltimore City. Additionally, these levels of income inequality have not changed much in the past 10 years.")
          )
        )
      )
    ),

    # ── Tab 2: Michelle ─────────────────────────────────────────────────────
    tabPanel(
      "Michelle",
      br(),
      sidebarLayout(
        sidebarPanel(
          width = 3,
          div(class = "section-label", "Question"),
          p("What geographic patterns exist in unemployment rates across the United States?",
            style = "font-size:12px; color:#6b7a90;"),
          hr(),
          div(class = "section-label", "Controls"),
          selectInput("michelle_variable", "Select a variable to map",
            choices = c(
              "Unemployment Rate"              = "unemp_rate",
              "Labor Force Participation Rate" = "lfpr",
              "Employment Rate"               = "emp_rate",
              "Not in Labor Force Rate"       = "nilf_rate"
            )
          ),
          sliderInput("michelle_year", "Select Year",
            min = 2015, max = 2022, value = 2022, step = 1, sep = "")
        ),
        mainPanel(
          width = 9,
          leafletOutput("michelle_map", height = "550px"),
          br(),
          fluidRow(
            column(4, wellPanel(strong("Highest"), textOutput("michelle_highest"))),
            column(4, wellPanel(strong("Lowest"),  textOutput("michelle_lowest"))),
            column(4, wellPanel(strong("US Median"), textOutput("michelle_med")))
          )
        )
      )
    ),

    # ── Tab 3: Ian ──────────────────────────────────────────────────────────
    tabPanel(
      "Ian",
      br(),
      sidebarLayout(
        sidebarPanel(
          width = 3,
          div(class = "section-label", "Question"),
          p("Ian's research question goes here.", style = "font-size:12px; color:#6b7a90;"),
          hr(),
          div(class = "section-label", "Controls"),
          p("Ian's controls go here.", style = "font-size:12px; color:#6b7a90;")
        ),
        mainPanel(
          width = 9,
          p("Ian's visualizations go here.", style = "font-size:13px; color:#5f5e5a;")
        )
      )
    ),

    # ── Tab 4: Summary ──────────────────────────────────────────────────────
    tabPanel(
      "Summary",
      br(),
      sidebarLayout(
        sidebarPanel(
          width = 3,
          div(class = "section-label", "Variables"),
          p("Select variables to summarize.", style = "font-size:12px; color:#6b7a90;"),
          hr(),
          actionButton("summary_go", "Show Summary", class = "btn-primary", style = "width:100%;")
        ),
        mainPanel(
          width = 9,
          uiOutput("summary_output")
        )
      )
    )
  )
)

server <- function(input, output, session) {

  rv <- reactiveValues(panel = NULL)

  # ── Load Gini data ─────────────────────────────────────────────────────────
  observeEvent(input$fetch, {
    renv_path <- path.expand("~/.Renviron")
    if (file.exists(renv_path)) readRenviron(renv_path)
    key_use <- Sys.getenv("CENSUS_API_KEY", unset = "")
    Sys.setenv(CENSUS_API_KEY = key_use)
    years <- seq.int(2010L, 2024L)
    showNotification("Fetching ACS from Census API…", id = "busy", duration = NULL)
    out <- purrr::map_dfr(years, function(yr) {
      fetch_gini_year("MD", yr, census_key = key_use)
    })
    removeNotification("busy")
    rv$panel <- out
    showNotification(
      sprintf("Loaded %d county-year rows (ACS 5-year end years %d–%d).",
              nrow(out), 2010L, 2024L),
      type = "message"
    )
  }, ignoreInit = TRUE)

  # ── Tab 1 reactives ────────────────────────────────────────────────────────
  map_data <- reactive({
    req(rv$panel)
    rv$panel |> dplyr::filter(year == input$year_map)
  })

  map_labeled <- reactive({
    d <- map_data()
    if (nrow(d) < 4) {
      d$lisa <- "Too few counties for LISA"
      return(d)
    }
    w <- build_county_weights(d)
    d$lisa <- lisa_quadrants(d$gini, w, alpha = 0.05)
    d
  })

  lisa_colours <- c(
    "High–High (high inequality cluster)" = "#b2182b",
    "Low–Low"                             = "#2166ac",
    "High–Low (outlier)"                  = "#d95f0e",
    "Low–High (outlier)"                  = "#7570b3",
    "Not significant"                     = "#525252",
    "No data"                             = "#bdbdbd",
    "Too few counties for LISA"           = "#636363",
    "Could not build spatial weights"     = "#636363",
    "Other"                               = "#737373"
  )

  output$map_or_placeholder <- renderUI({
    if (is.null(rv$panel)) {
      div(
        style = paste(
          "display: flex; align-items: center; justify-content: center;",
          "height: 520px; background-color: #d0d0d0; border-radius: 6px;"
        ),
        div(
          style = "text-align: center; color: #555; font-size: 1.15rem;",
          "Press the ", strong("Load Data"), " button to get started."
        )
      )
    } else {
      plotOutput("plot_map", height = 520)
    }
  })

  output$plot_map <- renderPlot({
    d <- map_labeled()
    d <- sf::st_transform(d, 4326)
    if (input$map_fill == "gini") {
      p <- ggplot2::ggplot(d) +
        ggplot2::geom_sf(
          ggplot2::aes(fill = gini, colour = lisa, alpha = gini_moe),
          linewidth = 0.25
        ) +
        ggplot2::scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
        ggplot2::scale_colour_manual(values = lisa_colours, na.value = "grey50", name = "LISA (border)") +
        ggplot2::scale_alpha(range = c(0.35, 1), name = "Gini MOE")
    } else {
      p <- ggplot2::ggplot(d) +
        ggplot2::geom_sf(
          ggplot2::aes(fill = lisa, alpha = gini, linewidth = gini_moe),
          colour = "gray15"
        ) +
        ggplot2::scale_fill_manual(
          values = c(
            "High–High (high inequality cluster)" = "#b2182b",
            "Low–Low"                             = "#2166ac",
            "High–Low (outlier)"                  = "#f4a582",
            "Low–High (outlier)"                  = "#92c5de",
            "Not significant"                     = "#e0e0e0",
            "No data"                             = "#bdbdbd",
            "Too few counties for LISA"           = "#636363",
            "Could not build spatial weights"     = "#636363",
            "Other"                               = "#cccccc"
          ),
          na.value = "grey90",
          name = "LISA fill"
        ) +
        ggplot2::scale_alpha(range = c(0.45, 1), name = "Gini") +
        ggplot2::scale_linewidth(range = c(0.05, 0.55), name = "Gini MOE")
    }
    p +
      ggplot2::coord_sf(crs = sf::st_crs(4326)) +
      ggplot2::labs(title = "County Inequality in Maryland") +
      ggplot2::theme_minimal(base_size = 12) +
      ggplot2::theme(
        panel.grid      = ggplot2::element_blank(),
        axis.text       = ggplot2::element_blank(),
        axis.title      = ggplot2::element_blank(),
        legend.position = "right"
      )
  })

  # ── Tab 2 (Michelle) reactives ─────────────────────────────────────────────
  michelle_data <- reactive({
    raw <- get_acs(
      geography = "county",
      variables = c(
        unemployed  = "B23025_005",
        labor_force = "B23025_002",
        employed    = "B23025_004",
        nilf        = "B23025_007",
        pop16       = "B23025_001"
      ),
      year     = input$michelle_year,
      survey   = "acs5",
      geometry = TRUE
    )

    wide <- raw |>
      sf::st_drop_geometry() |>
      tidyr::pivot_wider(
        id_cols     = c(GEOID, NAME),
        names_from  = variable,
        values_from = estimate
      )

    geom <- raw |>
      dplyr::select(GEOID) |>
      dplyr::group_by(GEOID) |>
      dplyr::slice(1) |>
      dplyr::ungroup()

    dplyr::left_join(geom, wide, by = "GEOID") |>
      dplyr::mutate(
        unemp_rate = round(unemployed  / labor_force * 100, 1),
        lfpr       = round(labor_force / pop16       * 100, 1),
        emp_rate   = round(employed    / labor_force * 100, 1),
        nilf_rate  = round(nilf        / pop16       * 100, 1)
      ) |>
      dplyr::filter(!is.na(unemp_rate), labor_force > 0)
  })

  michelle_var_label <- reactive({
    switch(input$michelle_variable,
      unemp_rate = "Unemployment Rate (By Quantile)",
      lfpr       = "Labor Force Participation Rate (By Quantile)",
      emp_rate   = "Employment Rate (By Quantile)",
      nilf_rate  = "Not in Labor Force Rate (By Quantile)"
    )
  })

  output$michelle_map <- renderLeaflet({
    df  <- michelle_data()
    col <- df[[input$michelle_variable]]
    pal <- colorQuantile("YlOrRd", col, n = 5)

    leaflet(df) |>
      addProviderTiles("CartoDB.Positron") |>
      setView(lng = -98.5, lat = 38.5, zoom = 4) |>
      addPolygons(
        fillColor    = ~pal(df[[input$michelle_variable]]),
        fillOpacity  = 0.7,
        color        = "white",
        weight       = 0.4,
        highlight    = highlightOptions(weight = 2, fillOpacity = 0.9),
        label        = ~paste0(NAME, ": ", df[[input$michelle_variable]], "%"),
        labelOptions = labelOptions(textsize = "13px")
      ) |>
      addLegend(
        position = "bottomright",
        pal      = pal,
        values   = col,
        title    = michelle_var_label()
      )
  })

  output$michelle_highest <- renderText({
    df  <- michelle_data() |> sf::st_drop_geometry()
    top <- df |> dplyr::slice_max(.data[[input$michelle_variable]], n = 1)
    paste0(top$NAME, " — ", top[[input$michelle_variable]], "%")
  })

  output$michelle_lowest <- renderText({
    df  <- michelle_data() |> sf::st_drop_geometry()
    bot <- df |> dplyr::slice_min(.data[[input$michelle_variable]], n = 1)
    paste0(bot$NAME, " — ", bot[[input$michelle_variable]], "%")
  })

  output$michelle_med <- renderText({
    paste0(median(michelle_data()[[input$michelle_variable]], na.rm = TRUE), "%")
  })

  # ── Tab 4: Summary — only fires on button press ────────────────────────────
  output$summary_output <- renderUI({

    # Nothing renders at all until the button has been clicked at least once
    req(input$summary_go)

    isolate({

      # ── Gini section ────────────────────────────────────────────────────
      gini_section <- if (is.null(rv$panel)) {
        p("No Gini data loaded yet. Press 'Load Data' on the Inequality Clusters tab first.",
          style = "font-size:13px; color:#888780;")
      } else {
        df <- rv$panel |>
          dplyr::select(NAME, year, gini, gini_moe) |>
          sf::st_drop_geometry()
        tagList(
          h4("Summary Statistics: Gini Index by County-Year (Maryland)",
             style = "font-size:15px; font-weight:600;"),
          renderTable({
            df |>
              dplyr::group_by(year) |>
              dplyr::summarise(
                N      = dplyr::n(),
                Mean   = round(mean(gini,   na.rm = TRUE), 4),
                Median = round(median(gini, na.rm = TRUE), 4),
                SD     = round(sd(gini,     na.rm = TRUE), 4),
                Min    = round(min(gini,    na.rm = TRUE), 4),
                Max    = round(max(gini,    na.rm = TRUE), 4),
                .groups = "drop"
              )
          }, striped = TRUE, hover = TRUE, bordered = TRUE)
        )
      }

      # ── Unemployment section ─────────────────────────────────────────────
      unemp_year <- isolate(input$michelle_year)

      unemp_section <- tryCatch({
        raw <- get_acs(
          geography = "county",
          variables = c(
            unemployed  = "B23025_005",
            labor_force = "B23025_002",
            employed    = "B23025_004",
            nilf        = "B23025_007",
            pop16       = "B23025_001"
          ),
          year   = unemp_year,
          survey = "acs5"
        )

        wide <- raw |>
          tidyr::pivot_wider(
            id_cols     = c(GEOID, NAME),
            names_from  = variable,
            values_from = estimate
          ) |>
          dplyr::mutate(
            unemp_rate = round(unemployed  / labor_force * 100, 1),
            lfpr       = round(labor_force / pop16       * 100, 1),
            emp_rate   = round(employed    / labor_force * 100, 1),
            nilf_rate  = round(nilf        / pop16       * 100, 1)
          ) |>
          dplyr::filter(!is.na(unemp_rate), labor_force > 0)

        summary_tbl <- data.frame(
          Variable = c(
            "Unemployment Rate (%)",
            "Labor Force Participation Rate (%)",
            "Employment Rate (%)",
            "Not in Labor Force Rate (%)"
          ),
          Mean   = c(round(mean(wide$unemp_rate, na.rm = TRUE), 1),
                     round(mean(wide$lfpr,       na.rm = TRUE), 1),
                     round(mean(wide$emp_rate,   na.rm = TRUE), 1),
                     round(mean(wide$nilf_rate,  na.rm = TRUE), 1)),
          Median = c(round(median(wide$unemp_rate, na.rm = TRUE), 1),
                     round(median(wide$lfpr,       na.rm = TRUE), 1),
                     round(median(wide$emp_rate,   na.rm = TRUE), 1),
                     round(median(wide$nilf_rate,  na.rm = TRUE), 1)),
          SD     = c(round(sd(wide$unemp_rate, na.rm = TRUE), 1),
                     round(sd(wide$lfpr,       na.rm = TRUE), 1),
                     round(sd(wide$emp_rate,   na.rm = TRUE), 1),
                     round(sd(wide$nilf_rate,  na.rm = TRUE), 1)),
          Min    = c(round(min(wide$unemp_rate, na.rm = TRUE), 1),
                     round(min(wide$lfpr,       na.rm = TRUE), 1),
                     round(min(wide$emp_rate,   na.rm = TRUE), 1),
                     round(min(wide$nilf_rate,  na.rm = TRUE), 1)),
          Max    = c(round(max(wide$unemp_rate, na.rm = TRUE), 1),
                     round(max(wide$lfpr,       na.rm = TRUE), 1),
                     round(max(wide$emp_rate,   na.rm = TRUE), 1),
                     round(max(wide$nilf_rate,  na.rm = TRUE), 1)),
          Counties = c(sum(!is.na(wide$unemp_rate)),
                       sum(!is.na(wide$lfpr)),
                       sum(!is.na(wide$emp_rate)),
                       sum(!is.na(wide$nilf_rate)))
        )

        tagList(
          hr(),
          h4(paste0("Summary Statistics: Unemployment & Labor Market (US Counties, ", unemp_year, ")"),
             style = "font-size:15px; font-weight:600;"),
          renderTable(summary_tbl, striped = TRUE, hover = TRUE, bordered = TRUE)
        )
      }, error = function(e) {
        p(paste("Could not load unemployment data:", conditionMessage(e)),
          style = "font-size:13px; color:#c0392b;")
      })

      tagList(gini_section, unemp_section)
    })
  })
}

shinyApp(ui, server)
```

## Overview

In this portion of the project, you will improve upon your previous EDA report to build an interactive dashboard using Shiny and create a set of slides to present your findings. Think of it as "something to show for" on your resume, so take pride in your work!

## Preparation

1. Study the feedback that your peers gave you to see if they make sense. If so, figure out how to incorporate it into the project. Note that each group member may have received comments from different peers, so gather them.  

2. A number of the EDA projects, though having good individual questions, lack an overall theme, i.e., how do you describe the goal of of your project in 1 or 2 sentences? If this is the case, I'd like to you to dig a bit deeper to **come up with a unifying theme**. Note that this may mean you have completely redo a plot or two. Incorporate this theme into 

  - the title of your Shiny app, and
  - the title of your slides.

## Instruction for your Shiny app

1. The app must have an well-organized interface with **multiple visualization tabs and one summary tab**.
2. Each group member is responsible for at least one tab, *not including* the summary tab.
3. For the visualization tab:

    - Each tab must be interactive
    - Each tab must correspond to a different plot type (okay if say tab 1 is `geom_point() + geom_smooth()` and tab 2 is `geom_boxplot + geom_smooth()`)
    - Each plot must include at least 4 variables
    - Each tab needs to have check boxes or radio button
    
4. For the data summary tab:

    - The tab must allow the user to choose some of the variables of interest and output the summary statistics, and bonus for outputting the distribution on top of that
    - you must include an **action button**: the user will only see the updated output **after** clicking the button.

## Instruction for your slides and presentation

1. The slides are clean and well organized, suitable for a typical scientific report
2. There must be an intro + background on the dataset, the theme of the project, and where you're coming from (i.e. motivation to study these questions)
3. This is a data visualization course, so in the main body of the slides, pictures speak louder than words!! Use concise bullet points only (or none at all). **Do not include paragraphs of writings**. When you present, speak over plots not the words. 
4. Each member should present their own plots. Otherwise, it's up to you who presents the intro, background, and conclusion. 
5. Bonus for a short demo of the Shiny app. 
6. The recorded presentation should be 10-15 minutes long. Feel free to make edits if you see necessary. Upload the final video to YouTube. Set it to private and include the shared link in your submission (see 1st bullet point in submission). 

## Submission

A `zip` file consisting of 

  - one `RMD` file and one `HTML` file, same as in the previous project, but **MAKE SURE TO INCLUDE YOUR YOUTUBE LINK AT THE VERY BEGINNING**
  - one `R` file for your Shiny app
  - one set of slides in **PDF** format (work in Google Slides then save to PDF)
  - one pdf file with the post-production report

## Grading structure

The precise rubric will be published via ELMS. In addition to the criteria from the previous projects, you will be graded on the functionality of your Shiny app and the quality of your presentation. 

## Group Structure 

> **Optional but recommended:** switching your role from the previous project for a more well-rounded experience.

> **Caution:** Please let me know of communication issues as soon as possible. You are not expected to pick up substantial slacks from your group members. 

| Role | Responsibilities |
|------|----------------|
| **Project Leader/Manager** | - Guides the overall direction of the EDA project <br> - Organizes workflow, deadlines, communication (with team members and the instructor), and file submission  |
|||
| **Data/Stat Lead** | - Performs data cleaning, transformations, and statistical analysis <br>- Provides technical support to others |
|||
| **Visualization/Communication Lead** | - Designs clear, well-labeled, and statistically informed figures <br>- Standardizes figures/texts and integrate them into report<br>|
|||
| **All Members** | - Each member is responsible for $\geq 1$ Shiny tabs and their own portion of the presentation and slides <br> - Each member is responsible for writing a group member review, to be submitted separately |

## Post-production

Each member is required to submit a report including an individual reflection and a group member evaluation. Both will be available only to the instructor/grader. Each report must include the following key information:

- Your roles and your plots (1 paragraph)
- Summary of your work (1 paragraph)
- Evaluation of each of your team member's work, i.e., what aspects are done well and what could be improved for each member. (1 paragraph for each member)
