---
title: "Geographic Patterns in Unemployment Rates"
author: "tidycensus Analysis"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidycensus)
library(shiny)
library(leaflet)
library(dplyr)
library(tidyr)
library(sf)
options(tigris_use_cache = TRUE)  # cache shapefiles so they don't re-download

# census_api_key("YOUR_KEY_HERE", install = TRUE)
# Restart R after running the line above — only needed once
```

## About this app

This app explores **Question 3: What geographic patterns exist in unemployment rates?**
using the American Community Survey (ACS) via the `tidycensus` package.

- **Geography**: County-level (national) or Census tract (by state)
- **Variables**: `B23025_005` (civilian unemployed) ÷ `B23025_002` (labor force)
- **Survey**: 5-year ACS — required for tract-level data, which is never available
  in the 1-year ACS ([Walker, Ch. 2](https://walker-data.com/census-r/an-introduction-to-tidycensus.html))

---

```{r ui}
fluidPage(

  # --- Controls ---
  wellPanel(
    fluidRow(
      column(4,
        selectInput("geography", "Geography level",
          choices = c(
            "County (national)" = "county",
            "Census tract (by state)" = "tract"
          )
        )
      ),
      column(4,
        # State selector — only meaningful for tracts
        # For counties we pull all US counties in one call
        conditionalPanel(
          condition = "input.geography == 'tract'",
          selectInput("state", "State (required for tracts)",
            choices = setNames(state.abb, state.name),
            selected = "MD"
          )
        )
      ),
      column(4,
        sliderInput("year", "ACS 5-year end year",
          min = 2015, max = 2022, value = 2022, step = 1, sep = ""
        )
      )
    )
  ),

  # --- Map ---
  leafletOutput("map", height = "520px"),

  br(),

  # --- Stats strip ---
  fluidRow(
    column(3, wellPanel(
      h5("Median rate"),
      textOutput("stat_median")
    )),
    column(3, wellPanel(
      h5("Highest county/tract"),
      textOutput("stat_max")
    )),
    column(3, wellPanel(
      h5("Lowest county/tract"),
      textOutput("stat_min")
    )),
    column(3, wellPanel(
      h5("Areas above 10%"),
      textOutput("stat_high")
    ))
  ),

  # --- Data table toggle ---
  checkboxInput("show_table", "Show underlying data table", value = FALSE),
  conditionalPanel(
    condition = "input.show_table == true",
    DT::dataTableOutput("data_table")
  )
)
```

```{r server}
# ── Reactive: fetch data from Census API ──────────────────────────────────────
map_data <- reactive({

  geo  <- input$geography
  yr   <- input$year
  st   <- if (geo == "tract") input$state else NULL

  # Fetch in default tidy (long) format, then pivot wide ourselves.
  # tidycensus v1.7+ with output="wide" names columns after the raw variable ID
  # (e.g. B23025_005E) not the custom name, so we pivot manually for safety.
  withProgress(message = "Fetching Census data...", value = 0.3, {

    df_long <- get_acs(
      geography = geo,
      variables = c(
        unemployed  = "B23025_005",   # Civilian unemployed
        labor_force = "B23025_002"    # In labor force
      ),
      state    = st,
      year     = yr,
      survey   = "acs5",             # 5-yr required for tract; Walker Ch.2 §2.2.1
      geometry = TRUE               # keep sf geometry on long-format result
    )

    setProgress(0.6)

    # Pivot to wide: one row per geography, columns: unemployed_E/M, labor_force_E/M
    # We use the named-variable labels that tidycensus puts in the `variable` column
    geo_col <- sf::st_geometry(df_long)   # preserve geometry before pivot

    df_wide <- df_long |>
      sf::st_drop_geometry() |>
      tidyr::pivot_wider(
        id_cols     = c(GEOID, NAME),
        names_from  = variable,
        values_from = c(estimate, moe),
        names_glue  = "{variable}_{.value}"   # → unemployed_estimate, etc.
      ) |>
      rename(
        unemployed_E  = unemployed_estimate,
        unemployed_M  = unemployed_moe,
        labor_force_E = labor_force_estimate,
        labor_force_M = labor_force_moe
      )

    # Re-attach geometry (one geometry row per GEOID after pivot)
    geom_df <- df_long |>
      select(GEOID) |>
      group_by(GEOID) |>
      slice(1) |>
      ungroup()

    df <- dplyr::left_join(geom_df, df_wide, by = "GEOID") |>
      mutate(
        unemp_rate = round(unemployed_E / labor_force_E * 100, 1),
        unemp_moe  = round(unemployed_M / labor_force_E * 100, 1),
        reliable   = !is.na(unemp_rate) & !is.na(unemp_moe) &
                       (unemp_moe < unemp_rate / 2)
      ) |>
      filter(!is.na(unemp_rate), labor_force_E > 0)

    setProgress(1)
    df
  })
})

# ── Render: Leaflet choropleth map ────────────────────────────────────────────
output$map <- renderLeaflet({

  df  <- map_data()

  # colorQuantile bins into equal-sized groups — better than colorNumeric
  # for right-skewed unemployment distributions (Walker Ch. 6 pattern)
  pal <- colorQuantile("YlOrRd", df$unemp_rate, n = 5, na.color = "#cccccc")

  labels <- sprintf(
    "<strong>%s</strong><br/>Unemployment: %s%%<br/>MOE: ±%s%%<br/>%s",
    df$NAME,
    df$unemp_rate,
    df$unemp_moe,
    ifelse(df$reliable, "", "<em style='color:#c0392b'>⚠ Estimate may be unreliable</em>")
  ) |> lapply(htmltools::HTML)

  leaflet(df) |>
    addProviderTiles("CartoDB.Positron") |>
    addPolygons(
      fillColor   = ~pal(unemp_rate),
      fillOpacity = 0.75,
      color       = "white",
      weight      = 0.5,
      highlight   = highlightOptions(
        weight      = 2,
        color       = "#444",
        fillOpacity = 0.9,
        bringToFront = TRUE
      ),
      label        = labels,
      labelOptions = labelOptions(
        style     = list("font-weight" = "normal", padding = "4px 8px"),
        textsize  = "13px",
        direction = "auto"
      )
    ) |>
    addLegend(
      position = "bottomright",
      pal      = pal,
      values   = ~unemp_rate,
      title    = paste("Unemployment<br>rate (%) —", input$year),
      opacity  = 0.85
    )
})

# ── Render: summary stats ─────────────────────────────────────────────────────
output$stat_median <- renderText({
  paste0(median(map_data()$unemp_rate, na.rm = TRUE), "%")
})

output$stat_max <- renderText({
  df  <- map_data()
  top <- df |> filter(unemp_rate == max(unemp_rate, na.rm = TRUE)) |> slice(1)
  paste0(top$NAME, " (", top$unemp_rate, "%)")
})

output$stat_min <- renderText({
  df  <- map_data()
  bot <- df |> filter(unemp_rate == min(unemp_rate, na.rm = TRUE)) |> slice(1)
  paste0(bot$NAME, " (", bot$unemp_rate, "%)")
})

output$stat_high <- renderText({
  n <- sum(map_data()$unemp_rate > 10, na.rm = TRUE)
  paste0(n, " area", ifelse(n == 1, "", "s"))
})

# ── Render: data table ────────────────────────────────────────────────────────
output$data_table <- DT::renderDataTable({
  map_data() |>
    sf::st_drop_geometry() |>
    select(
      GEOID,
      Name          = NAME,
      `Unemp. rate (%)` = unemp_rate,
      `MOE (±%)`    = unemp_moe,
      Reliable      = reliable,
      `Unemployed`  = unemployed_E,
      `Labor force` = labor_force_E
    ) |>
    arrange(desc(`Unemp. rate (%)`))
}, options = list(pageLength = 15, scrollX = TRUE))
```

---

### Notes on methodology

**Variable selection** — `B23025_005` and `B23025_002` come from ACS Table B23025
(Employment Status for the Population 16+). Use `load_variables()` to explore
alternatives:

```r
vars <- load_variables(2022, "acs5", cache = TRUE)
View(vars)  # filter label column for "unemployed"
```

**Margin of error** — ACS estimates carry a margin of error (MOE). For small
geographies like Census tracts, the MOE can exceed the estimate itself, making the
rate unreliable. This app flags those cases with a ⚠ warning in the map tooltip.

**Why 5-year ACS?** — The 1-year ACS only covers geographies with population ≥ 65,000,
so Census tracts are never available. The 5-year ACS (`survey = "acs5"`) is always
the right choice for sub-county analysis.

**Color scale** — `colorQuantile()` divides areas into equal-sized buckets (quintiles),
which prevents high-unemployment outliers from washing out color variation across
the rest of the map.
